---
title: "0 - data wrangling narrative"
author: "ER Deyle; Kayode; Janet O."
date: "2023-06-22"
output: html_document
---

This narrative describes the initial processing steps taken to fuse CDC data on human infections and NEON data on mosquito populations into a single data environment.

```{r setup, include=FALSE}
library(tidyverse)
library(sf)
```

File names for outputs

```{r}
fname_cases_by_county <- "./DATA/PROCESSED/CDC_WNV_cases_by_county.Rdata"
fname_mos_by_site <- "./DATA/PROCESSED/mos_total_counts.Rdata"
```

## Pulling data

NEON metadata gives us locations for mosquito etc.

```{r}
df_NEON_metadata_FULL <- read_csv("https://raw.githubusercontent.com/eco4cast/neon4cast-targets/main/NEON_Field_Site_Metadata_20220412.csv")
df_NEON_metadata <- df_NEON_metadata_FULL %>%
  select(field_site_id,field_site_county,field_site_state,field_latitude,field_longitude) %>%
  mutate(FullGeoName = paste0(field_site_state,", ",field_site_county))
```

## CDC Data

```{r}
if(file.exists(fname_cases_by_county)){
  load(fname_cases_by_county)
}else{
  source("_0_process_CDC.R")
}
```

## NEON mosquito abundance data

```{r}
if(file.exists(fname_mos_by_site)){
  load(fname_mos_by_site)
}else{
  source("_0_pull_trap_mosquito.R")
}
```


## Use NEON metadata to make an overlap; negotiating county codes

```{r pressure, echo=FALSE}
df_total_counts_by_county <- df_total_counts_by_site %>%
  left_join(df_NEON_metadata,by=c("siteID"="field_site_id")) %>%
  select(FullGeoName,everything())


df_cases_by_county <-  df_cases_by_county %>%
  filter(FullGeoName %in% unique(df_total_counts_by_county$FullGeoName)) %>%
  filter(Activity=="Human infections") %>%
  filter(Year >= 2014) %>%
  select(FullGeoName,Year,`Reported human cases`,`Neuroinvasive disease cases`) %>%
  complete(FullGeoName,Year,fill=list("Reported human cases"=0,"Neuroinvasive disease cases"=0))


# df_NEON_metadata
# df_cases_by_county
# df_NEON_metadata 

save(df_total_counts_by_county,df_cases_by_county,file="./DATA/PROCESSED/_0_merged_CDC_NEON.Rdata")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
