---
title: "Bird Diversity"
author: "Ayanna St. Rose"
date: "2023-06-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load all Libraries
```{r}
library(neonUtilities)
library(dplyr)
library(tidyverse)
library(neonUtilities)
library(multcomp)
library(janitor)
library(vegan)
```
#Source bird data from NEON data portal (Breeding landbird point counts DP1.10003.001)
```{r}
bird_data = loadByProduct(dpID = "DP1.10003.001", 
              site= c("BARR", "DSNY", "TOOL", "LAJA", "OAES", "WOOD", "NOGP", "TOOL", "DCFS", "JORN", "KONA", "BONA", "CLBJ", "DEJU", "GUAN", "HARV", "KONZ", "LENO", "MLBS","NIWO", "ONAQ", "ORNL", "OSBS", "SCBI", "SJER", "SOAP", "TALL", "UKFS", "UNDE", "WREF","YELL", "ABBY", "BART", "BLAN", "DELA",  "GRSM", "HEAL", "JERC", "MOAB", "TEAK", "PUUM", "SERC", "RMNP","STEI", "TREE"),
              startdate = "2013-01", enddate = "2022-12",
              check.size = FALSE)
```
#Aggregate site data
#Note that some years does not have data
#Extract expert taxonomic ID from data list
#Download brd_countdata table and bird_fielddata (metadata) from list of beetle field data
```{r}
bird_taxa = bird_data$brd_countdata

bird_fieldData = bird_data$brd_perpoint
```

#Merge taxonomic and field data by namedLocation
```{r}
bird_full_data = merge(bird_taxa, bird_fieldData,
                   by = c("eventID", "pointID", "uid", "plotID", "siteID", "domainID", "release",
                          "plotType", "startDate", "namedLocation", "publicationDate") ,
                   all = TRUE, check.duplicates = FALSE)
```

#Remove rows where taxon ID says NA
#This will have the same contents as the bird taxa dataframe,
#with added columns from field table
```{r}
bird_full_data = bird_full_data %>% drop_na(taxonID)
```

#Write bird presence data that includes taxonomic information and field information to csv
```{r}
write.csv(bird_full_data, "bird_full_data.csv", row.names = FALSE)
```

#Adjust start date format to year
```{r}
#Change format of collect dates to years
bird_full_data$startDate = data.frame(format(as.Date(bird_full_data$startDate, format = "%Y-%m-%d"), "%Y"))
```

###Goal:
#####Run species richness for each site for each year
#####Create data frame with different species for each site
#####Pseudoabundance is how many times the species is repeated (run for each site)
#####Count number of times a unique species occurs in each site
```{r}
#bird occurrence only sorted by siteID
#Quantify bird occurrence
#Group by specific epithet and site ID
bird_occ = as.matrix(bird_full_data[,c(5,9,14)])
bird_occ = data.frame(bird_occ)
colnames(bird_occ)[2] <- "year"

#Grouo by site then by specific epithet and year
bird_occ = bird_occ %>%
  tabyl(year, taxonID, siteID, sort = TRUE, head = TRUE)
```

#Prep the data for abundance
#Create for loop to calculate abundance in each site by year
#Show species distribution for each species for each site
```{r}
for (i in seq_along(bird_occ)){
  row.names(bird_occ[[i]]) <- (bird_occ[[i]][,1]) #Place row as header of column
  bird_occ[[i]] = bird_occ[[i]][,-c(1)] #Place delete year as a column
  bird_occ[[i]]$abundance = rowSums(bird_occ[[i]]) #get abundance by getting the number of species with 
}

abby = bird_occ$ABBY
```

#Pull out the richness column for each site
```{r}
bird_abun = lapply(bird_occ, function(x) x[,ncol(x)])

```

```{r}
bird = list_rbind(bird_occ, names_to = 'siteID')
bird = bird[, c(1, 570)]
```

```{r}
write.csv(bird, "bird_abundance.csv", row.names = FALSE)
```